---
title: "Project 4"
author: "Phil, Jess, Hannah"
date: "2023-10-01"
output: html_document
bibliography: BIOL3140.bib
---
```{r,include=FALSE}
library(tidyverse) #always
library(ape) #for evo analysis
library(nlme) #ditto
library(MuMIn) #for AIC
library(geiger) #for eco analysis
library(knitr) #for knitting options
```
# Introduction

The purpose of this project was to assess the relationship between anole hind limb-length and size (as measured by snout-vent length). Data collected in a study of allometry and ecomorphology of anoles (@hagey2017there) and tree data from a study of the phylogenetic relationships of anoles (@poe2017phylogenetic) were used to generate phylogenetically corrected models under BM and OU character evolution with respect to established covariates.

<center>
[![An anolis lizard](https://encrypted-tbn3.gstatic.com/images?q=tbn:ANd9GcQ3X6IkVdXlYV9zBSiFRqKy1_NDl2m7xTtIrQ3ohygmMH1yczwW)](https://encrypted-tbn3.gstatic.com/images?q=tbn:ANd9GcQ3X6IkVdXlYV9zBSiFRqKy1_NDl2m7xTtIrQ3ohygmMH1yczwW)
</center>

## Methods

Data from (@hagey2017there) was summarized in tables and log transformed for analysis. Using the nlme and ape packages in R and anole phylogenetic tree data from @poe2017phylogenetic, phylogenetic generalized least squares (PGLS) models were generated under BM and OU character evolution modes with and without Ecomorph as a covariate. Model fit was assessed with AIC scores using the MuMIN and geiger packages. 

### Results 

```{r, echo=FALSE, message=FALSE, label="load the data"}
anole <- read_csv("anole.dat.csv")
anole.eco <- read_csv("anole.eco.csv")
anole.tree <- read.tree("anole.tre")
```
```{r, echo=FALSE,message=FALSE, label= "join, filter, mutate to log"}
anole2 <- anole%>%
left_join(anole.eco)%>%
  filter(!Ecomorph%in%c("U","CH"))%>%
  na.omit()

anole.log <- anole2%>%
  mutate_at(c("SVL", "HTotal","PH","ArbPD"),log)
```
```{r, echo=FALSE,label= "PGLS models"}
### phylogenetic GLS models

#PGLS under BM, no ecomorph
pgls.BM1 <- gls(HTotal ~SVL, correlation = corBrownian(1,phy = anole.tree,form=~Species),data = anole.log, method = "ML")

#PGLS under BM, w ecomorph
pgls.BM2 <- gls(HTotal ~SVL * Ecomorph2, correlation = corBrownian(1,phy = anole.tree,form=~Species),data = anole.log, method = "ML")

#PGLS under OU, no ecomorph
pgls.OU1 <- gls(HTotal ~SVL, correlation = corMartins(0,phy = anole.tree,form=~Species),data = anole.log, method = "ML")

#PGLS under OU, w, ecomorph
pgls.OU2 <- gls(HTotal ~SVL * Ecomorph2, correlation = corMartins(0,phy = anole.tree,form=~Species),data = anole.log, method = "ML")
```
```{r,echo=FALSE,label="AIC"}
#AIC operations
anole.phylo.aic <- AICc(pgls.BM1,pgls.BM2,pgls.OU1,pgls.OU2)
anole.phylow.aicw <- aicw(anole.phylo.aic$AICc)
```
```{r, echo=FALSE, label="add residuals"}
#add phylo-corrected residuals
anole.log <- anole.log%>%
  mutate(phylo.res=residuals(pgls.BM2))
```{r, echo=FALSE, "adding AIC Table"}
anole.phylo.aic %>% 
  kable(caption = "Summary of PGLS Model AIC Scores(Key:BM1 = Hindlimb Length~Snout Vent Length no covariate, BM2 = Hindlimb Length~Snout Vent Length + Ecomorph covariate, OU1 = Hindlimb Length~Snout Vent Length no covariate, OU2 = Hindlimb Length~Snout Vent Length + Ecomorph covariate)")
```
PGLS model of hind-limb variation under BM character evolution and with Ecomorph as a covariate yielded the best fit. Model fit was visualized for each Ecomorph in the following graph: 
```{r, message = FALSE,echo=FALSE, label="Linear Regression of log-transformed Htotal vs SVL colored by covariateEcomorph"}
#Linear Regression of log-transformed Htotal vs SVL colored by covariateEcomorph
anole.log%>%
  ggplot(aes(HTotal,SVL,col=Ecomorph2))+geom_point()+geom_smooth(method="lm")
```









Residuals were plotted in the following graph:
```{r,echo=FALSE, label="plot residuals", fig.cap="Boxplot of phylogenetically corrected hindlimb residuals versus anole ecomorph"}
#plot residuals
p.eco.phylo <- anole.log%>%
ggplot(aes(x=Ecomorph2,y=phylo.res)) +geom_boxplot() +stat_summary(fun=mean, geom="point", size=3)
print(p.eco.phylo)
```


#### Discussion 


##### Author Contributions 

Group members completed the assignment separately and then came together to compare codes, answer questions, and create a final product. 

###### References 















